---
title: "Getting and Cleaning Data"
author: "qquito"
date: "27-08-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
```{r, include=FALSE}
library(dplyr)
setwd("../project3")
```

### Getting and Loading Data

\
Loads a vector with the activity codes in the "activity_labels.txt" file:
```{r}
activity.code <- c(WALKING=1, WALKING_UPSTAIRS=2, WALKING_DOWNSTAIRS=3, SITTING=4, STANDING=5, LAYING=6)
```

\
Loads the features to be used as column names, "features" is a data frame
```{r}
features <- read.csv("features.txt", sep=" ", header=FALSE)
```

\
This code, loads the "test" files and forms a new Data Frame with them. Here, test_data is a data frame to hold all the data from the test subjects, test_act is a data frame to hold the activity codes and test_sub is a data frame to hold all the subject codes. Finally we create a new data frame to store all the data in it and I called it "test_df".
```{r}
test_data <- read.fwf("./test/X_test.txt", widths=c(rep(16, 561)), header=FALSE)
test_act <- read.csv("./test/y_test.txt", header=FALSE)
test_sub <- read.csv("./test/subject_test.txt", header = FALSE)
test_df <- data.frame(test_sub, test_act, test_data)
```

\
This code, loads the "train" files and forms a new Data Frame with them. Here, train_data is a data frame to hold all the data from the train subjects, train_act is a data frame to hold the activity codes and train_sub is a data frame to hold all the subject codes in the train files. Finally we create a new data frame to store all the data in it and I called it "train_df"
```{r}
train_data <- read.fwf("./train/X_train.txt", widths=c(rep(16, 561)), header=FALSE)
train_act <- read.csv("./train/y_train.txt", header=FALSE)
train_sub <- read.csv("./train/subject_train.txt", header = FALSE)
train_df <- data.frame(train_sub, train_act, train_data)
```

\
Next, this code joins the two data frames (one of the objectives of the project [1]). The new data frame data shall be used from now on.
```{r}
data <- rbind(test_df, train_df)
```

\
This is optional, but here is a code to remove the data frames that are no longer being used.
```{r}
rm(test_data, test_act, test_sub, test_df)
rm(train_data, train_act, train_sub, train_df)
```

\
This line of code assigns the names of the columns to the data frame. Here we use the values loaded from the features file. (other of the objective of the project [4]). "features$V2" stores the list of names that we obtained before. Also, we add two column names to the subject code and the activity code.
```{r}
names(data) <- c("subject", "activity", features$V2)
```

\
Here I am extracting the unused measurements. The objective is to use only the mean and standard deviation (This is other of the objectives of the project [2]). Here is important to use the regular expression to get exactly the columns we want to use. In this code, R demonstrate its power.
```{r}
data <- data[names(data[grep("mean\\()|std\\()|subject|activity", names(data))])]
```

\
Here we assign the descriptive activity names to name the activities in the data set (Other of the objectives of the project [3]). One of the first lines of this script load the activity codes into a vector, and now we use it to update the data in the data frame.
```{r}
data$activity <- names(activity.code)[match(data$activity, activity.code)]
```

\
Finally, we create the tidy data set with the average of each variable for each activity and each subject (Objective [5]). It is interesting the use of the %>% operator, we first take the data frame and group_by it on subject and activity, and this way we just end summarizing it on all its variables.
```{r}
tidy_data <- data %>% group_by(subject, activity) %>% summarise_all(list(mean))
```

\
The final step... Writes the TXT file to the WD (File needed to submit the project)
```{r}
write.table(tidy_data, "./tidy_dataset.txt", row.names = FALSE)
```
